---
title: "Improving MCMC convergence diagnostic with a local version of R-hat: Known distributions on the chains"
author: "Khalil Leachouri, Théo Moins, Julyan Arbel, Stéphane Girard, Anne Dutfoy"
date: "6/18/2021"
output: 
  html_document: 
    self_contained: no
    toc: yes
---


```{r setup, include=FALSE}

knitr::opts_chunk$set(
  comment = NA, 
  cache = FALSE,
  fig.width = 12,
  fig.height = 8,
  message = FALSE,
  dev = c("png", "pdf")
)
options(width = 100)

# Working directory : localrhat
r_folder <- "../R/"
```


```{r, comment=NA, message=FALSE, warning=FALSE, results='hide', cache=FALSE}
library("EnvStats")
library('jmuOutlier')
library("ggplot2")
library("rstan")

devtools::load_all()

source(paste(r_folder, "import/monitornew.R", sep=""))
source(paste(r_folder, "import/r_star_monitor.R", sep=""))

N <- 500
reps <- 500
```


```{r, include=FALSE}
r_star_gbm <- function(chaines){
  dim_chaines <- dim(chaines)
  R_star_chains <- array(dim = c(dim_chaines[1], dim_chaines[2], 1))
  R_star_chains[,,1] <- chaines
  return (r_star(R_star_chains, method = "gbm"))
}

gen_chains <- function(M, N, rdists){
  return (array(c(rdists[[1]]((M-1)*N), rdists[[2]](N)), c(N,M)))
}

repetitions_R <- function(chains_func, r_func, r_names, reps=100){
  ncolumn <- length(r_func)
  R_values <- c()
  for (f in r_func){
    vect <- c()
    for (e in 1:reps){
      chaines <- chains_func()
      vect <- c(vect, f(chaines))
    }
    R_values = c(R_values, vect)
  }
  M <- matrix(data = R_values, ncol = ncolumn)
  colnames(M) <- r_names
  return (M)
}

colors <- c(rgb(0.1,0.6,0.1,0.7), rgb(0.1,0.1,0.7,0.7), rgb(0.5, 0, 0.5, 0.8), rgb(0.8, 0.1, 0.1, 0.8), rgb(1,0.6,0,0.8))

r_version <- c()
r_experiment <- c()
r_values <- c()

```

\begin{equation}
    \label{eq:R_theorique}
    R(x) = \sqrt{\frac{W(x)+B(x)}{W(x)}} = \sqrt{1 + \frac{\sum_{j=1}^m\sum_{k=j+1}^m \left(F_j(x)-F_k(x)\right)^2}{m\sum_{j=1}^m F_j(x)(1-F_j(x))}}.
\end{equation}

# Known distributions on the chains

### Example 2.a: Uniform distribution

As a first example we consider $m$ chains following centered uniform distributions, 
with the last having a different support than the other:
\begin{align*}
  F_1(x) &= \cdots = F_{m-1}(x) =  \frac{x}{2\sigma} + \frac{1}{2}, \quad \forall x \in \left[-\sigma; \sigma\right],\\
  \text{and}\quad F_m(x) &=  \frac{x}{2\sigma_m} + \frac{1}{2}, \quad \forall x \in \left[-\sigma_m; \sigma_m\right].
\end{align*}

In such a case the expression of $R(x)$ is explicit and can be compared to the estimator $\hat{R}(x)$:
to illustrate we choose $m=4$ chains, $\sigma = 3/4$ and $\sigma_m = 1$.

```{r uniform_r_x, warning=FALSE}
sigma <- 3/4
sigmaM <- 1
M <- 4

uniform_dists <- c(rep(c((function(q) punif(q, -sigma, sigma))), M-1),
                   (function(q) punif(q, -sigmaM, sigmaM)))

theoretical_r_values <- r_dist_values(npoints = N, xlim = c(-sigmaM, sigmaM),
                                      dists = uniform_dists)

uniform_rdists <- c((function(n) runif(n, -sigma, sigma)),
                    (function(n) runif(n, -sigmaM, sigmaM)))

chaines = gen_chains(M, N, uniform_rdists)
simulated_rhat = all_local_rhat(chaines, max_nb_points = "ALL")

xlabels <- c(1, 1.02, 1.04, 1.06, 1.08)

plot_local_r(chaines, simulated_rhat, theoretical_r_values, col=c(colors[3], colors[2]),
             xlabels = xlabels, plot_legend = FALSE,
             xlim = c(-1,1), ylim=c(1,1.09), title ="Uniform distributions")
```


We can see that $\hat{R}(x)$ (in violet) slightly overestimate $R(x)$, which makes it more conservative than
the population value. 

We replicate the experiment 500 times and study the histogram of replications of
different versions of $\hat{R}$:

```{r uniform_hist, warning=FALSE}
r_functions <- c(trad_rhat, rhat, rhat_infinity)
r_names <- c("R-hat",
             "Rank-R-hat",
             "R-hat-infinity")
r_colors <- c(colors[4], colors[5], colors[3])

R_matrix_unif <- repetitions_R(chains_func = (function() gen_chains(M, N, uniform_rdists)), 
                         r_func = r_functions, 
                         r_names = r_names, 
                         reps = reps)

xlabels <- c(1, 1.02, 1.04, 1.06, 1.08, 1.1)

plot_hist(R_matrix_unif, colors = r_colors, xlabels = xlabels,
          bin_size = 0.0033, lim_y_axis = reps)
```

As constructed initially by Vehtari et al. (2021), this example fool the original $\hat{R}$.
The rank version and $\hat{R}_\infty$ is on the contrary robust in this case.

```{r, include=FALSE}
r_version <- c(r_version, 
               rep(colnames(R_matrix_unif)[1], dim(R_matrix_unif)[1]),
               rep(colnames(R_matrix_unif)[2], dim(R_matrix_unif)[1]),
               rep(colnames(R_matrix_unif)[3], dim(R_matrix_unif)[1]))
r_experiment <- c(r_experiment, rep("1.Uniform", length(R_matrix_unif)))
r_values <- c(r_values, as.vector(R_matrix_unif))
```


### Example 2.b: Pareto distribution

Similarly, we check the second counter-example that fool the original $\hat{R}$:
chains with heavy-tails and different locations. To do so we consider Pareto 
distributions on the chains:
\begin{align*}
    F_1(x) &= \cdots = F_{m-1}(x) = 1 - \left({x}/{\eta}\right)^{-\alpha},
    \quad \forall x \in [\eta, +\infty),\\
    \text{and}\quad F_m(x) &=  1 - \left({x}/{\eta_m}\right)^{-\alpha},
    \quad \forall x \in [\eta_m, +\infty).
\end{align*}

As an example, we choose $\alpha = 0.8 \leq 1$ to ensure infinite moments and 
$(\eta, \eta_m) = (1, 1.5)$: 

```{r pareto_r_x, warning=FALSE}
eta <- 1
etaM <- 1.5
alpha <- 0.8
M <- 4

pareto_dists <- c(rep(c((function(q) ppareto(q, location = eta, shape = alpha))), M-1),
                  (function(q) ppareto(q, location = etaM, shape = alpha)))

theoretical_r_values <- r_dist_values(npoints = 1000, xlim = c(-eta, eta*8), 
                                      dists = pareto_dists)

pareto_rdists <- c((function(n) rpareto(n, location = eta, shape = alpha)),
                   (function(n) rpareto(n, location = etaM, shape = alpha)))

xlabels <- c(1, 1.02, 1.04, 1.06, 1.08)
chaines <- gen_chains(M, N, pareto_rdists)
simulated_rhat <- all_local_rhat(chaines, max_nb_points = "ALL")

plot_local_r(chaines, simulated_rhat, theoretical_r_values,
             plot_legend = FALSE, col=c(colors[3], colors[2]),
             xlabels = xlabels, xlim = c(1,8), ylim=c(1,1.09), title ="Pareto distributions")

```

The same observations as for the previous example can be done here: $\hat{R}_\infty$
overestimates $R_\infty$, whose value already allowed the diagnosis of a convergence issue.
We can verify the behaviour on replications: 

```{r pareto_hists, warning=FALSE}
r_functions <- c(trad_rhat, rhat, rhat_infinity)
r_names <- c("R-hat", "Rank-R-hat", "R-hat-infinity")
r_colors <- c(colors[4], colors[5], colors[3])

R_matrix_pareto <- repetitions_R(chains_func = (function() gen_chains(M, N, pareto_rdists)), 
                         r_func = r_functions, 
                         r_names = r_names, 
                         reps = reps)

xlabels = c(1, 1.02, 1.04, 1.06, 1.08)
plot_hist(R_matrix_pareto, colors = r_colors, xlabels = xlabels,
          plot_legend = FALSE, bin_size = 0.003, lim_y_axis = 500)
```

```{r, include=FALSE}
r_version <- c(r_version, 
               rep(colnames(R_matrix_pareto)[1], dim(R_matrix_pareto)[1]),
               rep(colnames(R_matrix_pareto)[2], dim(R_matrix_pareto)[1]),
               rep(colnames(R_matrix_pareto)[3], dim(R_matrix_pareto)[1]))  
r_experiment <- c(r_experiment, rep("2.Pareto", length(R_matrix_pareto)))
r_values <- c(r_values, as.vector(R_matrix_pareto))
```

## Counter examples for rank-$\hat{R}$

In this part, We will see 3 cases where rank-$\hat{R}$ doesn't manage to detect the convergence issue. 
These cases correspond to chains with different type of distributions, but with same mean and
mean over the median.

### Example 2.c: Uniform and Laplace distributions

We start by an example where the population version can be computed explicitly: 
if one chain is uniformly distributed $\mathcal{U}(-2\sigma, 2\sigma)$ and 
another one from a Laplace distribution $\mathcal{L}(0, \sigma)$, where $\sigma>0$,
then calculations leads to $R_\infty \approx 1.018$, which means convergence issue.


```{r laplace_uniform_x, warning=FALSE}
sigma <- 1/4
M <- 2


dists <- c((function(q) punif(q, -2*sigma, 2*sigma)),
           (function(q) plaplace(q, mean = 0, sd = sqrt(2)*sigma)))


theoretical_r_values <- r_dist_values(npoints = 1000, xlim = c(-1.5, 1.5), 
                                      dists = dists)


rdists <- c((function(n) runif(n, -2*sigma, 2*sigma)),
            (function(n) rlaplace(n, mean = 0, sd = sqrt(2)*sigma)))


chaines <- gen_chains(M, N, rdists)
simulated_rhat <- all_local_rhat(chaines, max_nb_points = "ALL")

xlabels <- c(1, 1.01, 1.02, 1.03, 1.04, 1.05)

plot_local_r(chaines, simulated_rhat, theoretical_r_values, threshold = 1.01,
             xlabels = xlabels, col = c(colors[3], colors[2]),
             xlim = c(-1.5,1.5), ylim=c(0.999,1.06), title ="Laplace/Uniform distributions")

```

Replications show that rank-$\hat{R}$ is fooled in the same way as $\hat{R}$ in this
example:


```{r laplace_uniform_hists, warning=FALSE}

r_functions <- c(trad_rhat, rhat, rhat_infinity)
r_names <- c("R-hat", "Rank-R-hat", "R-hat-infinity")
r_colors <- c(colors[4], colors[5], colors[3])

R_matrix_lapl_unif <- repetitions_R(chains_func = (function() gen_chains(M, N, rdists)), 
                         r_func = r_functions, 
                         r_names = r_names, 
                         reps = reps)

xlabels = c(1, 1.01, 1.02, 1.03, 1.04, 1.05)

plot_hist(R_matrix_lapl_unif, colors = r_colors, bin_size = 0.002,
          xlabels = xlabels, plot_legend = TRUE, threshold = 1.01,
          lim_y_axis = reps, vaxis_pos = 0.998)

```


```{r, include=FALSE}
r_version <- c(r_version, 
               rep(colnames(R_matrix_lapl_unif)[1], dim(R_matrix_lapl_unif)[1]),
               rep(colnames(R_matrix_lapl_unif)[2], dim(R_matrix_lapl_unif)[1]),
               rep(colnames(R_matrix_lapl_unif)[3], dim(R_matrix_lapl_unif)[1]))
r_experiment <- c(r_experiment, rep("3.Laplace/Uniform", length(R_matrix_lapl_unif)))
r_values <- c(r_values, as.vector(R_matrix_lapl_unif))
```

### Example 2.d: Uniform and Normal distributions

Similarly, others distributions are possible to build counter-examples. 
Consider $m-1$ normal chains $\mathcal{N}\left(0, \frac{\pi}{2}\sigma^2\right)$ and
for the last one a uniform $\mathcal{U}(-2\sigma, 2\sigma)$:

```{r gaussian_uniform_x, warning=FALSE}
sigma <- 1/2
M <- 2

dists <- c((function(q) pnorm(q, mean = 0, sd = sqrt(0.5*pi)*sigma)),
           (function(q) punif(q, -2*sigma, 2*sigma)))

theoretical_r_values <- r_dist_values(npoints = 1000, xlim = c(-2, 2), 
                                      dists = dists)

rdists <- c((function(n) rnorm(n, mean = 0, sd = sqrt(0.5*pi)*sigma)),
            (function(n) runif(n, -2*sigma, 2*sigma)))

chaines <- gen_chains(M, 2*N, rdists)
simulated_rhat <- all_local_rhat(chaines, max_nb_points = "ALL")

xlabels <- c(1, 1.005, 1.01, 1.015, 1.02)

plot_local_r(chaines, simulated_rhat, theoretical_r_values, 
             xlabels = xlabels, col=c(colors[3], colors[2]), threshold = 1.01,
             xlim = c(-2,2), ylim=c(0.999,1.025), title ="Gaussian/Uniform distributions")
```

Replications confirms that this construction are fooling the different versions
of $\hat{R}$ except $\hat{R}_\infty$.

```{r gaussian_uniform_hists, warning=FALSE}
r_functions <- c(trad_rhat, rhat, rhat_infinity)
r_names <- c("R-hat", "Rank-R-hat", "R-hat-infinity")
r_colors <- c(colors[4], colors[5], colors[3])

R_matrix_norm_unif <- repetitions_R(chains_func = (function() gen_chains(M, N, rdists)), 
                         r_func = r_functions, 
                         r_names = r_names, 
                         reps = reps)

xlabels = c(1, 1.01, 1.02, 1.03)
plot_hist(R_matrix_norm_unif, colors = r_colors, xlabels = xlabels,
          bin_size = 0.002, plot_legend = TRUE, threshold = 1.01,
          lim_y_axis = reps, vaxis_pos = 0.999)

```

```{r, include=FALSE}
r_version <- c(r_version, 
               rep(colnames(R_matrix_norm_unif)[1], dim(R_matrix_norm_unif)[1]),
               rep(colnames(R_matrix_norm_unif)[2], dim(R_matrix_norm_unif)[1]),
               rep(colnames(R_matrix_norm_unif)[3], dim(R_matrix_norm_unif)[1]))
r_experiment <- c(r_experiment, rep("4.Normal/Uniform", length(R_matrix_norm_unif)))
r_values <- c(r_values, as.vector(R_matrix_norm_unif))
```


### Example 2.e: Uniform and Exponential distributions

Finally, it is also possible to mix uniform distributions with exponential:

```{r exp_uniform_x, warning=FALSE}
M <- 4

min_unif <- 1-2*log(2)
max_unif <- 1+2*log(2)


dists <- c(rep(c((function(q) pexp(q))), M-1),
           (function(q) punif(q, min_unif, max_unif)))

theoretical_r_values <- r_dist_values(npoints = 500, xlim = c(-1, 6), 
                                      dists = dists)

rdists <- c((function(n) rexp(n)),
            (function(n) runif(n, min_unif, max_unif)))

chaines <- gen_chains(M, N, rdists)
simulated_rhat <- all_local_rhat(chaines, max_nb_points = "ALL")

xlabels <- c(1, 1.02, 1.04, 1.06, 1.08, 1.1)
plot_local_r(chaines, simulated_rhat, theoretical_r_values,
             xlabels = xlabels, col = c(colors[3], colors[2]),
             xlim = c(-1, 6), ylim=c(1,1.11), title ="Exponential/Uniform distributions")
```

```{r exp_uniform_hists, warning=FALSE}

r_functions <- c(trad_rhat, rhat, rhat_infinity)
r_names <- c("R-hat", "Rank-R-hat", "R-hat-infinity")
r_colors <- c(colors[4], colors[5], colors[3])

R_matrix_exp_unif <- repetitions_R(chains_func = (function() gen_chains(M, N, rdists)), 
                         r_func = r_functions, 
                         r_names = r_names, 
                         reps = reps)

plot_hist(R_matrix_exp_unif, colors = r_colors, bin_size = 0.005, 
          lim_y_axis = 500, vaxis_pos = 0.998)
```

```{r, include=FALSE}
r_version <- c(r_version, 
               rep(colnames(R_matrix_exp_unif)[1], dim(R_matrix_exp_unif)[1]),
               rep(colnames(R_matrix_exp_unif)[2], dim(R_matrix_exp_unif)[1]),
               rep(colnames(R_matrix_exp_unif)[3], dim(R_matrix_exp_unif)[1]))
r_experiment <- c(r_experiment, rep("5.Exp/Uniform", length(R_matrix_exp_unif)))
r_values <- c(r_values, as.vector(R_matrix_exp_unif))
```


## Chains with the same distributions

In this part we focus on the case where all the distributions are the same, to 
verify the behavior in the null hypothesis.

We start by looking at 4 chains uniformly distributed:

```{r same_uniform_x, warning=FALSE}
sup_unif <- 0.5
M <- 2

uniform_dists <- c(rep(c((function(q) punif(q, -sup_unif, sup_unif))), M))

theoretical_r_values <- r_dist_values(npoints = 1000, xlim = c(-sup_unif, sup_unif), 
                                      dists = uniform_dists)

uniform_rdists <- c((function(n) runif(n, -sup_unif, sup_unif)),
                    (function(n) runif(n, -sup_unif, sup_unif)))

chaines = gen_chains(M, N, uniform_rdists)
simulated_rhat = all_local_rhat(chaines, max_nb_points = "ALL")

plot_local_r(chaines, simulated_rhat, theoretical_r_values,
             col = c(colors[3], colors[2]),
             xlim = c(-0.6, 0.6), ylim=c(0.99,1.01), title ="Uniform distributions")

```

We can do the same analysis for pareto :

```{r same_pareto_x, warning=FALSE}
eta <- 1
alpha <- 0.8
M <- 2
N <- 500

pareto_dists <- c(rep(c((function(q) ppareto(q, location = eta, shape = alpha))), M))

theoretical_r_values <- r_dist_values(npoints = 1000, xlim = c(eta, eta*8), 
                                      dists = pareto_dists)

pareto_rdists <- c((function(n) rpareto(n, location = eta, shape = alpha)),
                   (function(n) rpareto(n, location = eta, shape = alpha)))

chaines = gen_chains(M, N, pareto_rdists)
simulated_rhat = all_local_rhat(chaines, max_nb_points = "ALL")

plot_local_r(chaines, simulated_rhat, theoretical_r_values,
             col = c(colors[3], colors[2]),
             xlim = c(eta, eta*8), ylim=c(0.99,1.01), title ="Pareto distributions")

```


```{r qqplot1, warning=FALSE}

M <- 2
N <- 500

R_matrix_same_unif <- repetitions_R(chains_func = (function() gen_chains(M, N, uniform_rdists)),
                                    r_func = c(rhat_infinity),
                                    r_names = c("R-hat-infinity"),
                                    reps = reps)

# R_matrix_unif <- N*(M-1)*(R_matrix_same_unif**2 - 1 + 1/N)
# R_matrix_unif <- N*M*(R_matrix_same_unif**2 - 1)

R_matrix_same_par <- repetitions_R(chains_func = (function() gen_chains(M, N, pareto_rdists)),
                                   r_func = c(rhat_infinity),
                                   r_names = c("R-hat-infinity"),
                                   reps = reps)

# R_matrix_par <- N*(M-1)*(R_matrix_same_par**2-1 +1/N)
# R_matrix_par <- N*M*(R_matrix_same_par**2-1)

dist1 <- sort((R_matrix_same_unif-mean(R_matrix_same_unif)))/sd(R_matrix_same_unif)
dist2 <- sort((R_matrix_same_par-mean(R_matrix_same_par)))/sd(R_matrix_same_par)

# pdf(file = "/scratch/tmoins/Documents/Code_rhat/figure_article/qqplot1.pdf", width = 8, height = 7)
ggplot(mapping = aes(x = dist1, y = dist2)) + 
  geom_point() +
  geom_abline(aes(slope = 1, intercept = 0), linetype = 2) +
  xlab("") + ylab("Pareto distribution") + 
  xlim(-2, 5) + ylim(-2, 5) +
  theme(axis.text=element_text(size=20), axis.title=element_text(size=24,face="bold"))
# dev.off()
```


```{r same_uniform_hists, warning=FALSE}

M <- 4
N <- 500

R_matrix_same_unif <- repetitions_R(chains_func = (function() gen_chains(M, N, uniform_rdists)),
                                    r_func = c(rhat_infinity),
                                    r_names = c("R-hat-infinity"),
                                    reps = reps)

# R_matrix_unif <- N*(M-1)*(R_matrix_same_unif**2 - 1 + 1/N)
# R_matrix_unif <- N*M*(R_matrix_same_unif**2 - 1)

R_matrix_same_par <- repetitions_R(chains_func = (function() gen_chains(M, N, pareto_rdists)),
                                   r_func = c(rhat_infinity),
                                   r_names = c("R-hat-infinity"),
                                   reps = reps)

# R_matrix_par <- N*(M-1)*(R_matrix_same_par**2-1 +1/N)
# R_matrix_par <- N*M*(R_matrix_same_par**2-1)

dist1 <- sort((R_matrix_same_unif-mean(R_matrix_same_unif))/sd(R_matrix_same_unif))
dist2 <- sort((R_matrix_same_par-mean(R_matrix_same_par))/sd(R_matrix_same_par))

# pdf(file = "/scratch/tmoins/Documents/Code_rhat/figure_article/qqplot2.pdf", width = 8, height = 7)
ggplot(mapping = aes(x = dist1, y = dist2)) + 
  geom_point() +
  geom_abline(aes(slope = 1, intercept = 0), linetype = 2) +
  xlab("Uniform distribution") + ylab("") +
  xlim(-2, 5) + ylim(-2, 5) +
  theme(axis.text=element_text(size=20), axis.title=element_text(size=24,face="bold"))
# dev.off()
```


```{r same_uniform_hists, warning=FALSE}

M <- 8
N <- 500

R_matrix_same_unif <- repetitions_R(chains_func = (function() gen_chains(M, N, uniform_rdists)),
                                    r_func = c(rhat_infinity),
                                    r_names = c("R-hat-infinity"),
                                    reps = reps)

# R_matrix_unif <- N*(M-1)*(R_matrix_same_unif**2 - 1 + 1/N)
# R_matrix_unif <- N*M*(R_matrix_same_unif**2 - 1)

R_matrix_same_par <- repetitions_R(chains_func = (function() gen_chains(M, N, pareto_rdists)),
                                   r_func = c(rhat_infinity),
                                   r_names = c("R-hat-infinity"),
                                   reps = reps)

# R_matrix_par <- N*(M-1)*(R_matrix_same_par**2-1 +1/N)
# R_matrix_par <- N*M*(R_matrix_same_par**2-1)

dist1 <- sort((R_matrix_same_unif-mean(R_matrix_same_unif))/sd(R_matrix_same_unif))
dist2 <- sort((R_matrix_same_par-mean(R_matrix_same_par))/sd(R_matrix_same_par))

# pdf(file = "/scratch/tmoins/Documents/Code_rhat/figure_article/qqplot3.pdf", width = 8, height = 7)
ggplot(mapping = aes(x = dist1, y = dist2)) + 
  geom_point() +
  geom_abline(aes(slope = 1, intercept = 0), linetype = 2) +
  xlab("") + ylab("") + 
  xlim(-2, 5) + ylim(-2, 5) +
  theme(axis.text=element_text(size=20), axis.title=element_text(size=24,face="bold"))
# dev.off()
```


```{r same_uniform_hists, warning=FALSE}
M <- 2
N <- 200

r_functions <- c(rhat_infinity)
r_names <- c("R-hat-infinity")
r_colors <- c(colors[3])

R_matrix_same_unif <- repetitions_R(chains_func = (function() gen_chains(M, N, uniform_rdists)), 
                                    r_func = r_functions, 
                                    r_names = r_names, 
                                    reps = reps)

# pdf(file = "/scratch/tmoins/Documents/Code_rhat/figure_article/same_unif_1.pdf", width = 8, height = 7)
xlabels = c(1, 1.01, 1.02, 1.03, 1.04)
plot_hist(R_matrix_same_unif, colors = r_colors, plot_threshold = T,
          threshold = 1.01, xlabels = xlabels, bin_size = 0.0020, 
          lim_y_axis = 200, vaxis_pos = 1, plot_legend = F)
# dev.off()
```


```{r same_uniform_hists, warning=FALSE}
M <- 4
N <- 100

r_functions <- c(rhat_infinity)
r_names <- c("R-hat-infinity")
r_colors <- c(colors[3])

R_matrix_same_unif <- repetitions_R(chains_func = (function() gen_chains(M, N, uniform_rdists)), 
                                    r_func = r_functions, 
                                    r_names = r_names, 
                                    reps = reps)

# pdf(file = "/scratch/tmoins/Documents/Code_rhat/figure_article/same_unif_2.pdf", width = 8, height = 7)
xlabels = c(1, 1.01, 1.02, 1.03, 1.04)
plot_hist(R_matrix_same_unif, colors = r_colors, plot_threshold = F,
          xlabels = xlabels, bin_size = 0.002, 
          lim_y_axis = 200, vaxis_pos = 1, plot_legend = F)
# dev.off()
```


```{r same_uniform_hists, warning=FALSE}
M <- 8
N <- 50

r_functions <- c(rhat_infinity)
r_names <- c("R-hat-infinity")
r_colors <- c(colors[3])

R_matrix_same_unif <- repetitions_R(chains_func = (function() gen_chains(M, N, uniform_rdists)), 
                                    r_func = r_functions, 
                                    r_names = r_names, 
                                    reps = reps)

# pdf(file = "/scratch/tmoins/Documents/Code_rhat/figure_article/same_unif_3.pdf", width = 8, height = 7)
xlabels = c(1, 1.01, 1.02, 1.03, 1.04)
plot_hist(R_matrix_same_unif, colors = r_colors, plot_threshold = F,
          threshold = 1.03, xlabels = xlabels, bin_size = 0.002, 
          lim_y_axis = 200, vaxis_pos = 1)
# dev.off()
```


```{r, include=FALSE}
get_r_lim_infinity <- function(M, N, alpha, 
                               reps = 500, N_estimation = 500){
  
  R_matrix_same_unif <- repetitions_R(chains_func = (function() gen_chains(M, N_estimation, uniform_rdists)),
                                      r_func = c(rhat_infinity),
                                      r_names = c("R-hat-infinity"),
                                      reps = reps)
  # q <- quantile(N_estimation*(M-1)*(R_matrix_same_unif**2 - 1 + 1/N_estimation), 
                                 # probs = c(1-alpha))
  q <- quantile(N_estimation*M*(R_matrix_same_unif**2 - 1), 
                                 probs = c(1-alpha))
  return (sqrt(1+ q/(N*M)))
}

get_alpha_infinity <- function(M, N, rlim, reps = 500,
                               N_estimation = 500){
  
  R_matrix_same_unif <- repetitions_R(chains_func = (function() gen_chains(M, N_estimation, uniform_rdists)),
                                      r_func = c(rhat_infinity),
                                      r_names = c("R-hat-infinity"),
                                      reps = reps)
  # empirical_dist <- ecdf(N_estimation*(M-1)*(R_matrix_same_unif**2 - 1 + 1/N_estimation))  
  empirical_dist <- ecdf(N_estimation*M*(R_matrix_same_unif**2 - 1))
  return (empirical_dist(N*M*(rlim**2-1)))
}

m_range <- c(2, 4, 8) #, 10, 15)
alpha_range <- c(0.001, 0.005, 0.01, 0.05, 0.1)
r_lim_tab <- c()
for (m in m_range){
  r_row <- c(m)
  for (alpha in alpha_range){
    r_row <- c(r_row, get_r_lim_infinity(M = m, N = 400/m, alpha = alpha))
  }
  r_lim_tab <- rbind(r_lim_tab, r_row)
}
colnames(r_lim_tab) <- c("m", alpha_range)
r_lim_tab
```

```{r qqplot1, warning=FALSE}

M <- 2
N <- 500

R_matrix_same_unif <- repetitions_R(chains_func = (function() gen_chains(M, N, uniform_rdists)),
                                    r_func = c(trad_rhat),
                                    r_names = c("R-hat"),
                                    reps = reps)

# R_matrix_unif <- N*(M-1)*(R_matrix_same_unif**2 - 1 + 1/N)
# R_matrix_unif <- N*M*(R_matrix_same_unif**2 - 1)

R_matrix_same_par <- repetitions_R(chains_func = (function() gen_chains(M, N, pareto_rdists)),
                                   r_func = c(trad_rhat),
                                   r_names = c("R-hat"),
                                   reps = reps)

# R_matrix_par <- N*(M-1)*(R_matrix_same_par**2-1 +1/N)
# R_matrix_par <- N*M*(R_matrix_same_par**2-1)

dist1 <- sort((R_matrix_same_unif-mean(R_matrix_same_unif)))/sd(R_matrix_same_unif)
dist2 <- sort((R_matrix_same_par-mean(R_matrix_same_par)))/sd(R_matrix_same_par)

# pdf(file = "/scratch/tmoins/Documents/Code_rhat/figure_article/qqplot1.pdf", width = 8, height = 7)
ggplot(mapping = aes(x = dist1, y = dist2)) + 
  geom_point() +
  geom_abline(aes(slope = 1, intercept = 0), linetype = 2) +
  xlab("") + ylab("Pareto distribution") + 
  xlim(-2, 5) + ylim(-2, 5) +
  theme(axis.text=element_text(size=20), axis.title=element_text(size=24,face="bold"))
# dev.off()
```





```{r same_uniform_hists, warning=FALSE}
# 
# M <- 4
# N <- 500
# 
# R_matrix_same_par <- repetitions_R(chains_func = (function() gen_chains(M, N, pareto_rdists)),
#                                      r_func = c((function(chains) local_rhat(2, chains))),
#                                      r_names = c("R-hat-infinity"),
#                                      reps = reps)
# 
# R_matrix <- N*(M-1)*(R_matrix_same_par**2-1 +1/N)
# bin_size <- 1
# lim_y_axis <- 0.5
# 
# 
# plot_hist_chi_square <- function(){
#   par(mar=c(4,5.5,2,2))
# 
#   xlim <- c(min(R_matrix), max(R_matrix))
# 
#   means = round(seq(from = min(R_matrix), to = max(R_matrix), length.out = 6), digits = 2)
# 
#   h <- ceiling((max(R_matrix[,1])-min(R_matrix[,1])) / bin_size)+1
# 
#   b <- seq(min(R_matrix[,1]),  max(R_matrix[,1]), length.out = h)
# 
#   hist(R_matrix[,1], col = colors[3], breaks = b,
#        xlim=xlim, ylim = c(0,lim_y_axis), add=FALSE,
#        xlab="", yaxt="n", xaxt="n", ylab="", main="", prob=TRUE)
#   means <- c(means, round(mean(R_matrix[,1]), digits=2))
# 
#   axis(2, cex.axis = 2, las=1, pos = 0, lwd = 4)
#   axis(1, cex.axis = 2, pos=0, padj=0.5, lwd = 4)
# 
#   xfit<-seq(0,max(R_matrix[,1]),length=100)
#   yfit<-dchisq(xfit, df= M-1)
#   lines(xfit, yfit, col=colors[1], lwd=5)
# }
# 
# # mean(R_matrix[,1])
# plot_hist_chi_square()
```



```{r, include=FALSE}
r_version2 <- c()
r_experiment2 <- c()
r_values2 <- c()

r_version2 <- c(r_version2, 
               rep(colnames(R_matrix_same_unif)[1], dim(R_matrix_same_unif)[1]),
               rep(colnames(R_matrix_same_unif)[2], dim(R_matrix_same_unif)[1]),
               rep(colnames(R_matrix_same_unif)[3], dim(R_matrix_same_unif)[1]))
r_experiment2 <- c(r_experiment2, rep("6. Same Uniform", length(R_matrix_same_unif)))
r_values2 <- c(r_values2, as.vector(R_matrix_same_unif))

r_version2 <- c(r_version2, 
               rep(colnames(R_matrix_same_par)[1], dim(R_matrix_same_par)[1]),
               rep(colnames(R_matrix_same_par)[2], dim(R_matrix_same_par)[1]),
               rep(colnames(R_matrix_same_par)[3], dim(R_matrix_same_par)[1]))
r_experiment2 <- c(r_experiment2, rep("7.Same Pareto", length(R_matrix_same_par)))
r_values2 <- c(r_values2, as.vector(R_matrix_same_par))
```

```{r r_star, warning=FALSE}
# reps <- 50
# 
# R_matrix_r_star = repetitions_R(chains_func = (function() gen_chains(M, N, pareto_rdists)), 
#                          r_func = c(r_star_gbm), 
#                          r_names = c("R-star"), 
#                          reps = reps)
# 
# plot_hist(R_matrix_r_star, colors = colors[4], bin_size = 0.08, 
#           lim_y_axis = 30, vaxis_pos = 0.78)

```
# Summary

```{r}
data=data.frame(r_version, r_experiment,  r_values)

ggplot(data, aes(x=r_experiment, y=r_values, fill=r_version)) +
    geom_boxplot() + 
    geom_hline(yintercept=1.01, linetype="dashed", 
               color = "black", size=0.5) +
    xlab("Experiments") + ylab("Values")
```



```{r}
data2=data.frame(r_version2, r_experiment2,  r_values2)

ggplot(data2, aes(x=r_experiment2, y=r_values2, fill=r_version2)) +
    geom_boxplot() + 
    geom_hline(yintercept=1.01, linetype="dashed", 
               color = "black", size=0.5) +
    xlab("Experiments") + ylab("Values")
```

