---
title: "Improving MCMC convergence diagnostic with a local version of R-hat: Empirical models"
author: "Khalil Leachouri, Théo Moins, Julyan Arbel, Stéphane Girard, Anne Dutfoy"
date: "6/18/2021"
output: 
  html_document: 
    self_contained: no
    toc: yes
---


```{r setup, include=FALSE}

knitr::opts_chunk$set(
  comment = NA, 
  cache = FALSE,
  fig.width = 14,
  message = FALSE,
  dev = c("png", "pdf")
)
options(width = 100)

# Working directory : localrhat
stan_folder <- "./stan_files/"
r_folder <- "../R/"
```


```{r, comment=NA, message=FALSE, warning=FALSE, results='hide', cache=FALSE}
library("EnvStats")
library('jmuOutlier')
library("ggplot2")
library("rstan")

source(paste(r_folder, "import/monitornew.R", sep=""))
source(paste(r_folder, "import/r_star_monitor.R", sep=""))

devtools::load_all()

N <- 500
reps <- 50
```


```{r, include=FALSE}
r_star_gbm <- function(chaines){
  dim_chaines <- dim(chaines)
  R_star_chains <- array(dim = c(dim_chaines[1], dim_chaines[2], 1))
  R_star_chains[,,1] <- chaines
  return (r_star(R_star_chains, method = "gbm"))
}


repetitions_R <- function(chains_func, r_func, r_names, 
                          reps=100, saved_model=FALSE){
  ncolumn <- length(r_func)
  R_values <- c()
  for (f in r_func){
    vect <- c()
    for (e in 1:reps){
      if (saved_model){
        chaines <- chains_func(e)
      } else {
        chaines <- chains_func() 
      }
      vect <- c(vect, f(chaines))
    }
    R_values = c(R_values, vect)
  }
  M <- matrix(data = R_values, ncol = ncolumn)
  colnames(M) <- r_names
  return (M)
}

colors <- c(rgb(0.1,0.6,0.1,0.7), rgb(0.1,0.1,0.7,0.7), rgb(0.5, 0, 0.5, 0.8), rgb(0.8, 0.1, 0.1, 0.8), rgb(1,0.6,0,0.8))

r_version <- c()
r_experiment <- c()
r_values <- c()

```

# Experiments when the population value is unknown

## Autoregressive model


```{r}
autoregressive <- function(n, rho, sigma){
  chain <- c()
  pre <- 0
  eps <- rnorm(n, 0, sigma)
  for (t in 1:n){
    suiv <- rho*pre+eps[t]
    chain <- c(chain, suiv)
    pre <- suiv
  }
  return(chain)
}

gen_autoreg_chains <- function(M, N, rho, sigma, sigmaM){
  return (array(c(replicate(M-1, autoregressive(N, rho, sigma)), 
                  autoregressive(N, rho, sigmaM)), c(N,M)))
}

```

## Plot of R-hat(x)

```{r}

M <- 4
sigma <- 1
sigmaM <- 2
rho <- 1/2


chaines = gen_autoreg_chains(M, N, rho, sigma, sigmaM)
simulated_rhat = all_local_rhat(chaines, max_nb_points = "ALL")

# pdf(file = "/scratch/tmoins/Documents/Code_rhat/figure_article/autoreg.pdf", width = 12, height = 7)
plot_local_r(chaines, simulated_rhat, col=c(colors[3], colors[2]),
             xlim = c(-6,6), ylim=c(1,1.07), title ="")
# dev.off()
```

## Repetitions: comparison of different R-hat

```{r uniform_hist, warning=FALSE}

r_functions <- c(trad_rhat, rhat, rhat_infinity)
r_names <- c("R-hat", "Rank-R-hat", "R-hat-infinity")
r_colors <- c(colors[4], colors[5], colors[3])

R_matrix_autoreg <- repetitions_R(
  chains_func = (function() gen_autoreg_chains(M, N, rho, sigma, sigmaM)), 
                         r_func = r_functions, 
                         r_names = r_names, 
                         reps = reps)

# pdf(file = "/scratch/tmoins/Documents/Code_rhat/figure_article/hist_autoreg.pdf", width = 12, height = 7)
plot_hist(R_matrix_autoreg, colors = r_colors, 
          xlabels = c(1, 1.02, 1.04, 1.06, 1.08, 1.1),
          bin_size = 0.004, lim_y_axis = 80)
# dev.off()
```

```{r, include=FALSE}
r_version <- c(r_version, 
               rep(colnames(R_matrix_autoreg)[1], dim(R_matrix_autoreg)[1]),
               rep(colnames(R_matrix_autoreg)[2], dim(R_matrix_autoreg)[1]),
               rep(colnames(R_matrix_autoreg)[3], dim(R_matrix_autoreg)[1]))
r_experiment <- c(r_experiment, rep("1.Autoregressive", length(R_matrix_autoreg)))
r_values <- c(r_values, as.vector(R_matrix_autoreg))
```


# Cauchy

In this section, we will study two parameterization of Cauchy using HMC:
- Nominal parameterization which is a direct one.
- Alternative parameterization which is a scale mixture of Gaussians.

To study these two parametrizations we will use rank plots which illustrate a mixing problem if they are not uniform. First we will look for an element of the vector for which our $\hat R$ is one of the highest. Here we want rather to study the tail of the distribution than the bulk because of the thick tail of the Cauchy distribution. Finally, we will use the rank plots of this element found to know if we have a mixing problem or not.


## Nominal parameterization

Model :

```{r}
writeLines(readLines(paste(stan_folder, "cauchy_nom.stan", sep="")))
```


```{r, warning=FALSE}
# fit_nom <- stan(file = paste(stan_folder, "cauchy_nom.stan", sep=""), seed = 7878, refresh = 0)
# saveRDS(fit_nom, paste(stan_folder, "cauchy_nom.rds", sep=""))
fit_nom <- readRDS(paste(stan_folder, "cauchy_fits/cauchy_nom.rds", sep=""))

mon <- monitor(fit_nom)

which_max_rhat <- which.max(mon[1:50, 'Rhat'])
max_rhat <- max(mon[1:50, 'Rhat'])

subset(mon, Rhat == max_rhat)
```
### Plot of R-hat(x)

```{r, warning=FALSE}
chains <- as.array(fit_nom)
dim_chains <- dim(chains)

rhat_inf <- c()

for (i in 1:dim_chains[3]) {
  chains_i <- chains[, , i]
  rhat_inf_i <- rhat_infinity(chains_i)
  rhat_inf <- c(rhat_inf, rhat_inf_i)
}

mon[["Rhat_inf"]] <- rhat_inf
which_max_rhat_inf <- which.max(mon[1:50, 'Rhat_inf'])
  
simulated_rhat = all_local_rhat(chains[ , , which_max_rhat_inf], max_nb_points = "ALL")

xlabels = 1:10
# pdf(file = "/scratch/tmoins/Documents/Code_rhat/figure_article/cauchy.pdf", width = 12, height = 7)
plot_local_r(chains[ , , which_max_rhat_inf], simulated_rhat,
             col=c(colors[3], colors[2]), xlabels = xlabels,
             xlim = c(-50,200), ylim=c(1,10), title ="")
# dev.off()
```


### Repetitions: comparison of different R-hat


```{r, warning=FALSE}

gen_diverging_cauchy <- function(i){
  fit_nom <- stan(file = paste(stan_folder, "cauchy_nom.stan", sep=""), refresh = 0)
  saveRDS(fit_nom, paste(stan_folder, "cauchy_fits/cauchy_nom_", i, ".rds", sep=""))
  # fit_nom <- readRDS(paste(stan_folder, "cauchy_fits/cauchy_nom_", i, ".rds", sep=""))
  
  
  mon <- monitor(fit_nom)
  chains <- as.array(fit_nom)
  dim_chains <- dim(chains)
  
  rhat_inf <- c()
  
  for (i in 1:dim_chains[3]) {
    chains_i <- chains[, , i]
    rhat_inf_i <- rhat_infinity(chains_i)
    rhat_inf <- c(rhat_inf, rhat_inf_i)
  }
  mon[["Rhat_inf"]] <- rhat_inf
  which_max_rhat_inf <- which.max(mon[1:50, 'Rhat_inf'])
  
  return (chains[, , which_max_rhat_inf])
}

r_functions <- c(trad_rhat, rhat, rhat_infinity)
r_names <- c("R-hat", "Rank-R-hat", "R-hat-infinity")
r_colors <- c(colors[4], colors[5], colors[3])

R_matrix_cauchy <- repetitions_R(chains_func = gen_diverging_cauchy, 
                                 r_func = r_functions, 
                                 r_names = r_names, 
                                 reps = reps, saved_model = TRUE)

xlabels = c(1, 1.1, 1.2, 1.3, 1.4, 1.5, 1.6)
# pdf(file = "/scratch/tmoins/Documents/Code_rhat/figure_article/hist_cauchy.pdf", width = 12, height = 7)
plot_hist(R_matrix_cauchy, colors = r_colors, xlabels = xlabels,
          bin_size = 0.02, lim_y_axis = reps/2)
# dev.off()
```


```{r, include=FALSE}
r_version <- c(r_version, 
               rep(colnames(R_matrix_cauchy)[1], dim(R_matrix_cauchy)[1]),
               rep(colnames(R_matrix_cauchy)[2], dim(R_matrix_cauchy)[1]),
               rep(colnames(R_matrix_cauchy)[3], dim(R_matrix_cauchy)[1]))
r_experiment <- c(r_experiment, rep("2.Cauchy", length(R_matrix_cauchy)))
r_values <- c(r_values, as.vector(R_matrix_cauchy))
```


## Alternative parameterization

Model :

```{r}
writeLines(readLines(paste(stan_folder, "cauchy_alt_1.stan", sep="")))
```

```{r, warning=FALSE}
fit_alt1 <- stan(file = paste(stan_folder, "cauchy_alt_1.stan", sep=""), seed = 7878, refresh = 0)
saveRDS(fit_alt1, paste(stan_folder, "cauchy_fits/cauchy_alt.rds", sep=""))
# fit_alt1 <- readRDS(paste(stan_folder, "cauchy_fits/cauchy_alt.rds", sep=""))
mon <- monitor(fit_alt1)


which_max_rhat <- which.max(mon[101:150, 'Rhat'])
max_rhat <- max(mon[101:150, 'Rhat'])
subset(mon, Rhat == max_rhat)

```

### Plot of R-hat(x)

```{r, warning=FALSE}
chains <- as.array(fit_alt1)
dim_chains <- dim(chains)

rhat_inf <- c()

for (i in 1:dim_chains[3]) {
  chains_i <- chains[, , i]
  rhat_inf_i <- rhat_infinity(chains_i)
  rhat_inf <- c(rhat_inf, rhat_inf_i)
}

mon[["Rhat_inf"]] <- rhat_inf
which_max_rhat_inf <- which.max(mon[101:150, 'Rhat_inf'])
  
simulated_rhat = all_local_rhat(chains[ , , which_max_rhat_inf], max_nb_points = "ALL")

xlabels = c(1, 1.005, 1.01)
# pdf(file = "/scratch/tmoins/Documents/Code_rhat/figure_article/cauchy_alt.pdf", width = 12, height = 7)
plot_local_r(chains[, , which_max_rhat], simulated_rhat, 
             col=c(colors[3], colors[2]), xlabels = xlabels,
             xlim = c(-4,4), ylim=c(0.9995,1.012), title ="")
# dev.off()
```

### Repetitions: comparison of different R-hat


```{r, warning=FALSE}

gen_converging_cauchy <- function(i){
  fit_nom <- stan(file = paste(stan_folder, "cauchy_alt_1.stan", sep=""), refresh = 0)
  saveRDS(fit_alt1, paste(stan_folder, "cauchy_fits/cauchy_alt_", i, ".rds", sep=""))
  # fit_alt1 <- readRDS(paste(stan_folder, "cauchy_fits/cauchy_alt_", i, ".rds", sep=""))
  
  mon <- monitor(fit_nom)
  chains <- as.array(fit_nom)
  dim_chains <- dim(chains)
  
  rhat_inf <- c()
  
  for (i in 1:dim_chains[3]) {
    chains_i <- chains[, , i]
    rhat_inf_i <- rhat_infinity(chains_i)
    rhat_inf <- c(rhat_inf, rhat_inf_i)
  }
  mon[["Rhat_inf"]] <- rhat_inf
  which_max_rhat_inf <- which.max(mon[101:150, 'Rhat_inf'])
  
  return (chains[, , which_max_rhat_inf])
}

r_functions <- c(trad_rhat, rhat, rhat_infinity)
r_names <- c("R-hat", "Rank-R-hat", "R-hat-infinity")
r_colors <- c(colors[4], colors[5], colors[3])

R_matrix_cauchy2 <- repetitions_R(chains_func = gen_converging_cauchy, 
                                  r_func = r_functions, 
                                  r_names = r_names, 
                                  reps = reps,
                                  saved_model = TRUE)

xlabels = c(1, 1.005, 1.01, 1.015, 1.02, 1.025, 1.03)
# pdf(file = "/scratch/tmoins/Documents/Code_rhat/figure_article/hist_cauchy_alt.pdf", width = 12, height = 7)
plot_hist(R_matrix_cauchy2, colors = r_colors, xlabels = xlabels,
          bin_size = 0.0007, lim_y_axis = reps)
# dev.off()
```

```{r, include=FALSE}
r_version <- c(r_version, 
               rep(colnames(R_matrix_cauchy2)[1], dim(R_matrix_cauchy2)[1]),
               rep(colnames(R_matrix_cauchy2)[2], dim(R_matrix_cauchy2)[1]),
               rep(colnames(R_matrix_cauchy2)[3], dim(R_matrix_cauchy2)[1]))
r_experiment <- c(r_experiment, rep("3.Cauchy2", length(R_matrix_cauchy2)))
r_values <- c(r_values, as.vector(R_matrix_cauchy2))
```

# Hierarchical model: Eight Schools

Model :

```{r}
writeLines(readLines(paste(stan_folder, "eight_schools_cp.stan", sep="")))
```

## Centered Eight Schools model

We will proceed in the same way as the case of the Cauchy distribution concerning the rank plots, but this time we will focus on the parameter $\tau$. And our choice is due to the fact that the distribution becomes more centralized on the population mean when faced with the small values of the population standard deviation $\tau$, which makes it difficult to explore this distribution by MCMC chains.

```{r, warning=FALSE}
eight_schools <- read_rdump(paste(stan_folder, "eight_schools.data.R", sep=""))
fit_cp <- stan(
  file = paste(stan_folder, "eight_schools_cp.stan", sep=""), data = eight_schools,
  iter = 2000, chains = 4, seed = 120, refresh = 0,
  control = list(adapt_delta = 0.95)
)
saveRDS(fit_cp, paste(stan_folder, "eight_schools_fits/eight_schools_cp.rds", sep=""))
# fit_cp <- readRDS(paste(stan_folder, "eight_schools_fits/eight_schools_cp.rds", sep=""))

mon <- monitor(fit_cp)
print(mon)
```
### Plot of R-hat(x)

```{r, warning=FALSE}
chains <- as.array(fit_cp)
dim_chains <- dim(chains)

simulated_rhat = all_local_rhat(chains[, , "tau"], max_nb_points = "ALL")

xlabels = c(1, 1.01, 1.02, 1.03, 1.04)

# pdf(file = "/scratch/tmoins/Documents/Code_rhat/figure_article/tau_cp.pdf", width = 12, height = 7)
plot_local_r(chains[, , "tau"], simulated_rhat, 
             col=c(colors[3], colors[2]), xlabels = xlabels,
             xlim = c(0,25), ylim=c(0.999,1.04), title ="")
# dev.off()
```

### Repetitions: comparison of different R-hat


```{r, warning=FALSE}

gen_centered_eight_school <- function(i){
  fit_cp <- stan(
    file = paste(stan_folder, "eight_schools_cp.stan", sep=""), data = eight_schools,
    iter = 2000, chains = 4, refresh = 0,
    control = list(adapt_delta = 0.95)
  )
  saveRDS(fit_cp, paste(stan_folder, "eight_schools_fits/eight_schools_cp_", i, ".rds", sep=""))
  # fit_cp <- readRDS(paste(stan_folder, "eight_schools_fits/eight_schools_cp_", i, ".rds", sep=""))
  
  mon <- monitor(fit_cp)
  chains <- as.array(fit_cp)

  return (chains[, , "tau"])
}

r_functions <- c(trad_rhat, rhat, rhat_infinity)
r_names <- c("R-hat", "Rank-R-hat", "R-hat-infinity")
r_colors <- c(colors[4], colors[5], colors[3])

R_matrix_8schools <- repetitions_R(chains_func = gen_centered_eight_school, 
                                   r_func = r_functions, 
                                   r_names = r_names, 
                                   reps = reps)

xlabels = c(1, 1.02, 1.04, 1.06, 1.08, 1.1, 1.12, 1.14)
# pdf(file = "/scratch/tmoins/Documents/Code_rhat/figure_article/hist_tau_cp.pdf", width = 12, height = 7)
plot_hist(R_matrix_8schools, colors = r_colors, xlabels = xlabels,
          bin_size = 0.005, lim_y_axis = 25)
# dev.off()
```
```{r, include=FALSE}
r_version <- c(r_version, 
               rep(colnames(R_matrix_8schools)[1], dim(R_matrix_8schools)[1]),
               rep(colnames(R_matrix_8schools)[2], dim(R_matrix_8schools)[1]),
               rep(colnames(R_matrix_8schools)[3], dim(R_matrix_8schools)[1]))
r_experiment <- c(r_experiment, rep("4.8Schools", length(R_matrix_8schools)))
r_values <- c(r_values, as.vector(R_matrix_8schools))
```

### Multivariate version on the likelihood

```{r, warning=FALSE}
chains <- as.array(fit_cp)
dim_chains <- dim(chains)

simulated_rhat = all_local_rhat(chains[, , "lp__"], max_nb_points = "ALL")

# xlabels = c(1, 1.005, 1.01, 1.015, 1.02)
plot_local_r(chains[, , "lp__"], simulated_rhat, 
             col=c(colors[3], colors[2])
             xlim = c(-7,7), ylim=c(0.999,1.02), title ="Likelihood")
```


## Non-centered Eight Schools model

Model :

```{r}
writeLines(readLines(paste(stan_folder, "eight_schools_ncp.stan", sep="")))
```

```{r, warning=FALSE}
fit_ncp2 <- stan(
  file = paste(stan_folder, "eight_schools_ncp.stan", sep=""), data = eight_schools,
  iter = 2000, chains = 4, seed = 483892929, refresh = 0,
  control = list(adapt_delta = 0.95)
)
saveRDS(fit_ncp2, paste(stan_folder, "eight_schools_fits/eight_schools_ncp.rds", sep=""))
# fit_ncp2 <- readRDS(paste(stan_folder, "eight_schools_fits/eight_schools_ncp.rds", sep=""))


mon <- monitor(fit_ncp2)
print(mon)
```

### Plot of R-hat(x)

```{r, warning=FALSE}
chains <- as.array(fit_ncp2)
dim_chains <- dim(chains)

simulated_rhat = all_local_rhat(chains[, , "tau"], max_nb_points = "ALL")

xlabels = c(1, 1.005, 1.01)
# pdf(file = "/scratch/tmoins/Documents/Code_rhat/figure_article/tau_ncp.pdf", width = 12, height = 7)
plot_local_r(chains[, , "tau"], simulated_rhat, 
             col=c(colors[3], colors[2]), xlabels = xlabels,
             xlim = c(0,25), ylim=c(0.999,1.011), title ="")
# dev.off()
```

### Repetitions: comparison of different R-hat

```{r, warning=FALSE}

gen_ncp_eight_school <- function(i){
  fit_ncp2 <- stan(
    file = paste(stan_folder, "eight_schools_ncp.stan", sep=""), data = eight_schools,
    iter = 2000, chains = 4, refresh = 0,
    control = list(adapt_delta = 0.95)
  )
  saveRDS(fit_ncp2, paste(stan_folder, "eight_schools_fits/eight_schools_ncp_", i, ".rds", sep=""))
  # fit_ncp2 <- readRDS(paste(stan_folder, "eight_schools_fits/eight_schools_ncp_", i, ".rds", sep=""))
  mon <- monitor(fit_ncp2)
  chains <- as.array(fit_ncp2)

  return (chains[, , "tau"])
}

r_functions <- c(trad_rhat, rhat, rhat_infinity)
r_names <- c("R-hat", "Rank-R-hat", "R-hat-infinity")
r_colors <- c(colors[4], colors[5], colors[3])

R_matrix_8schools2 <- repetitions_R(chains_func = gen_ncp_eight_school, 
                         r_func = r_functions, 
                         r_names = r_names, 
                         reps = reps,
                         saved_model = TRUE)

xlabels = c(1, 1.005, 1.01, 1.015, 1.02)
# pdf(file = "/scratch/tmoins/Documents/Code_rhat/figure_article/hist_tau_ncp.pdf", width = 12, height = 7)
plot_hist(R_matrix_8schools2, colors = r_colors, xlabels = xlabels,
          bin_size = 0.0007, lim_y_axis = reps)
# dev.off()
```
```{r, include=FALSE}
r_version <- c(r_version, 
               rep(colnames(R_matrix_8schools2)[1], dim(R_matrix_8schools2)[1]),
               rep(colnames(R_matrix_8schools2)[2], dim(R_matrix_8schools2)[1]),
               rep(colnames(R_matrix_8schools2)[3], dim(R_matrix_8schools2)[1]))
r_experiment <- c(r_experiment, rep("5.8Schools2", length(R_matrix_8schools2)))
r_values <- c(r_values, as.vector(R_matrix_8schools2))
```

### Multivariate version on the likelihood

```{r, warning=FALSE}
chains <- as.array(fit_ncp2)
dim_chains <- dim(chains)

simulated_rhat = all_local_rhat(chains[, , "lp__"], max_nb_points = "ALL")

xlabels = c(1, 1.005, 1.01, 1.015, 1.02)
plot_local_r(chains[, , "lp__"], simulated_rhat, xlabels = xlabels,
             xlim = c(-7,0), ylim=c(0.999,1.02), title ="Likelihood")
```

```{r}
data=data.frame(r_version, r_experiment,  r_values)

ggplot(data, aes(x=r_experiment, y=r_values, fill=r_version)) +
    geom_boxplot() + 
    geom_hline(yintercept=1.01, linetype="dashed", 
               color = "black", size=0.5) +
    xlab("Experiments") + ylab("Values") + ylim(1, 1.3)
```

