---
title: "Improving MCMC convergence diagnostic with a local version of R-hat: Multivariate simulations"
author: "Khalil Leachouri, Théo Moins, Julyan Arbel, Stéphane Girard, Anne Dutfoy"
date: "6/18/2021"
output: 
  html_document: 
    self_contained: no
    toc: yes
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(
  comment = NA, 
  cache = FALSE,
  fig.width = 12,
  fig.eight = 8,
  dev = c("png", "pdf")
)
options(width = 100)

# Working directory : localrhat
r_folder <- "../R/"
stan_folder <- "./stan_files/"
```

```{r, comment=NA, message=FALSE, warning=FALSE, results='hide', cache=FALSE}
library("EnvStats")
library('jmuOutlier')
library("rstan")
library("MASS") # For Multivariate Normal Distribution
library("stableGR") # For the R-hat version of Vats and Knudson (2018)
library("coda") # For an implementation of the multivariate R-hat of Brooks and Gelman (1998)
library("rWishart") # Wishart distribution
library("mcmc") # for logit data

library("ggplot2")
library("ggridges")


devtools::load_all()

source(paste(r_folder, "import/monitornew.R", sep=""))
source(paste(r_folder, "import/r_star_monitor.R", sep=""))

```

# Multivariate examples

In the multivariate case when $\boldsymbol{\theta}=(\theta_1, \ldots, \theta_d)\in{\mathbb R}^d$, 
we suggest to apply $\hat{R}$ on a multivariate indicator variable 
$I_{\boldsymbol{x}}^{(j)} = \mathbb{I}\{\theta_1^{(j)} \leq x_1, \ldots, \theta_d^{(j)} \leq x_d\}$ 
for any $\boldsymbol{x} = (x_1, \ldots, x_d) \in \mathbb{R}^d$ in two steps:

\begin{enumerate}
    \item Compute the univariate $\hat{R}_{\infty,p}$ separately on each of the 
    coordinates $p\in\{1,\dots,d\}$.
    \item If $\hat{R}_{\infty,p} < 1.01$ for all $p\in\{1,\dots,d\}$, then all 
    margins are deemed to have converged and to be identically distributed.
    Compute the multivariate $\hat{R}_\infty$ to check the convergence of the 
    dependence structure.
\end{enumerate}

Thus we recommend to compute the multivariate version only in the case where all the
univariate one doesn't detect any convergence issue, this is why in the part we will
focus only of cases where the margins are the same (for example uniform).


```{r, include=FALSE}
r_star_gbm <- function(chaines){
  R_star_chains = aperm(chaines, c(1, 3, 2))
  return (r_star(R_star_chains, method = "gbm"))
}

brooks_multivariate_rhat <- function(chaines){
  R_star_chains = aperm(chaines, c(1, 3, 2))
  return (r_hat_multivariate(R_star_chains, split_chains = F))
}

coda_multivariate_rhat <- function(chaines){
  chaines_list = list()
  for (i in 1:dim(chaines)[3]){
    mcmc_ch <- as.mcmc(chaines[,,i])
    chaines_list[[i]] <- as.mcmc(chaines[,,i])
  }
  mcmc_chains = do.call(mcmc.list, list(chaines_list))
  return (gelman.diag(mcmc_chains)$mpsrf)
}


vats_multivariate_rhat <- function(chaines){
  chaines_list = list()
  for (i in 1:dim(chaines)[3]){
    mcmc_ch <- as.mcmc(chaines[,,i])
    chaines_list[[i]] <- chaines[,,i]
  }
  return (stable.GR(chaines_list)$mpsrf)
}

repetitions_R <- function(chains_func, r_func, r_names, 
                          reps=100, saved_model=FALSE){
  ncolumn <- length(r_func)
  R_values <- c()
  for (f in r_func){
    vect <- c()
    for (e in 1:reps){
      if (saved_model){
        chaines <- chains_func(e)
      } else {
        chaines <- chains_func() 
      }
      vect <- c(vect, f(chaines))
    }
    R_values = c(R_values, vect)
  }
  M <- matrix(data = R_values, ncol = ncolumn)
  colnames(M) <- r_names
  return (M)
}

colors <- c(rgb(0.1,0.6,0.1,0.7), rgb(0.1,0.1,0.7,0.7), rgb(0.5, 0, 0.5, 0.8), rgb(0.8, 0.1, 0.1, 0.8), rgb(1,0.6,0,0.8))

```

## Choice of threshold

```{r, warning=FALSE}

gen_mvnormal_chains <- function(M, N, d, S=NULL){
  if (is.null(S)){
    S = generate_sigma_matrix(d)
  }
  return (array(c(mvrnorm((M-1)*N, mu = rep(0, d), Sigma = diag(d)), 
                  mvrnorm(N, mu = rep(0, d), Sigma = S)), 
                c(N,d,M)))
}

d <- 3
M <- 2
N <- 200
reps <- 200
# sig_matrix <- (1-rho) * diag(2) + matrix(rho, nrow=2, ncol=2)


r_functions = c(rhat_infinity)
r_names = c("R-hat-infinity")
r_colors = c(colors[3], colors[4])

R_matrix = repetitions_R(chains_func = (function() gen_mvnormal_chains(M, N, d, S = diag(d))), 
                         r_func = r_functions, 
                         r_names = r_names, 
                         reps = reps)

xlabels = c(1, 1.01, 1.02, 1.03, 1.04)
plot_hist(R_matrix, colors = r_colors, plot_threshold = T,
          threshold = 1.01, xlabels = xlabels, bin_size = 0.0020, 
          lim_y_axis = 100, vaxis_pos = 1)

```



```{r, warning=FALSE}

d <- 3
M <- 4
N <- 100
reps <- 200
# sig_matrix <- (1-rho) * diag(2) + matrix(rho, nrow=2, ncol=2)


r_functions = c(rhat_infinity)
r_names = c("R-hat-infinity")
r_colors = c(colors[3], colors[4])

R_matrix = repetitions_R(chains_func = (function() gen_mvnormal_chains(M, N, d, S = diag(d))), 
                         r_func = r_functions, 
                         r_names = r_names, 
                         reps = reps)

xlabels = c(1, 1.01, 1.02, 1.03, 1.04)
plot_hist(R_matrix, colors = r_colors, plot_threshold = F,
          xlabels = xlabels, bin_size = 0.0018, 
          lim_y_axis = 100, vaxis_pos = 1)

```

```{r, warning=FALSE}

d <- 3
M <- 8
N <- 50
reps <- 200
# sig_matrix <- (1-rho) * diag(2) + matrix(rho, nrow=2, ncol=2)


r_functions = c(rhat_infinity)
r_names = c("R-hat-infinity")
r_colors = c(colors[3], colors[4])

R_matrix = repetitions_R(chains_func = (function() gen_mvnormal_chains(M, N, d, S = diag(d))), 
                         r_func = r_functions, 
                         r_names = r_names, 
                         reps = reps)

xlabels = c(1, 1.01, 1.02, 1.03, 1.04, 1.05)
plot_hist(R_matrix, colors = r_colors, plot_threshold = F,
          threshold = 1.03, xlabels = xlabels, bin_size = 0.002, 
          lim_y_axis = 200, vaxis_pos = 1)

```


```{r, include=FALSE}
get_r_lim_infinity <- function(M, N, alpha, d,
                               reps = 100, N_estimation = 500){
  
  R_matrix <- repetitions_R(chains_func = (function() gen_mvnormal_chains(M, N_estimation, d, S = diag(d))), 
                                      r_func = c(rhat_infinity),
                                      r_names = c("R-hat-infinity"),
                                      reps = reps)
  q <- quantile(N_estimation*M*(R_matrix**2 - 1), 
                                 probs = c(1-alpha))
  return (sqrt(1+ q/(N*M)))
}

d_range <- c(2, 3, 4, 5)
m_range <- c(2, 4, 8) #, 10, 15)
alpha_range <- c(0.01, 0.05, 0.1) #c(0.001, 0.005, 0.01, 0.05, 0.1)

for (d in d_range){
  r_lim_tab_d <- c()
  for (m in m_range){
    r_row <- c(m)
    for (alpha in alpha_range){
      r_row <- c(r_row, get_r_lim_infinity(M = m, N = 400/m, alpha = alpha, d = d))
    }
    r_lim_tab_d <- rbind(r_lim_tab_d, r_row)
  }
  colnames(r_lim_tab_d) <- c("m", alpha_range)
  print(r_lim_tab_d)
}
```






```{r, include=FALSE}
get_r_lim_infinity <- function(M, N, alpha, d,
                               reps = 100, N_estimation = 200){
  
  R_matrix <- repetitions_R(chains_func = (function() gen_mvnormal_chains(M, N_estimation, d, S = diag(d))), 
                                      r_func = c(rhat_infinity_max_directions),
                                      r_names = c("max-R-hat"),
                                      reps = reps)
  q <- quantile(N_estimation*M*(R_matrix**2 - 1), 
                                 probs = c(1-alpha))
  return (sqrt(1+ q/(N*M)))
}

d_range <- c(2, 3, 4, 5)
m_range <- c(2, 4, 8) #, 10, 15)
alpha_range <- c(0.01, 0.05, 0.1) #c(0.001, 0.005, 0.01, 0.05, 0.1)

for (d in d_range){
  r_lim_tab_d <- c()
  for (m in m_range){
    r_row <- c(m)
    for (alpha in alpha_range){
      r_row <- c(r_row, get_r_lim_infinity(M = m, N = 400/m, alpha = alpha, 
                                           d = d))
    }
    r_lim_tab_d <- rbind(r_lim_tab_d, r_row)
  }
  colnames(r_lim_tab_d) <- c("m", alpha_range)
  print(r_lim_tab_d)
}
```






## Bivariate Case: Normal distributions

In this section, we focus on two-dimensional parameters on which we aim at 
diagnose convergence of the corresponding MCMC (with any number of chains).
This restriction allows to navigate easily between the positive dependence 
and the negative one of the two components.

In particular, we consider bivariate normal densities with zero means, and compare 
$m-1$ chains with identity covariance matrix, and one with non-zero off-diagonal
elements $\rho \in (-1,1)$.

We start by define the function that will allows us to generate $m$ chains of 
size $n$ with this configuration :

```{r, warning=FALSE}
gen_bvnormal_chains <- function(M, N, rho){
  sig_matrix <- (1-rho) * diag(2) + matrix(rho, nrow=2, ncol=2)
  return (array(c(mvrnorm((M-1)*N, mu = rep(0, 2), Sigma = diag(2)), 
                  mvrnorm(N, mu = rep(0, 2), Sigma = sig_matrix)), c(N,2,M)))
}
```


As an example, we consider the case with $m=4$ chains of size $n=100$, and 
$\rho = 0.9$ on the last chain. 
This choice of $n$ relatively small is due to the fact the elements in the chain
are i.i.d here, so $n$ can be seen as an effective sample size for a given chain
here.

We simulate 100 experiments, which result as 100 draws of the different $\hat{R}$.

Histogram of the corresponding values of $\hat{R}_\infty$ and the multivariate
$\hat{R}$ of Brooks and Gelman (1998) :

```{r bivariate_hists, warning=FALSE}
rho <- 0.9
M <- 2
N <- 200
reps <- 100

r_functions = c(rhat_infinity, brooks_multivariate_rhat)
r_names = c("R-hat-infinity", "Brooks Multivariate R-hat")
r_colors = c(colors[3], colors[4])

R_matrix = repetitions_R(chains_func = (function() gen_bvnormal_chains(M, N, rho)), 
                         r_func = r_functions, 
                         r_names = r_names, 
                         reps = reps)

xlabels = c(1, 1.02, 1.04, 1.06, 1.08, 1.1, 1.12)
# pdf(file = "/scratch/tmoins/Documents/Code_rhat/figure_article/hist_bivariate.pdf", width = 12, height = 7)
plot_hist(R_matrix, colors = r_colors, bin_size = 0.0045, 
          lim_y_axis = 30, vaxis_pos = 0.995,
          xlabels = xlabels, plot_threshold = F)
# dev.off()
```

In comparison, the value of $R^*$ of Lambert and Vehtari (2020) seems to detect 
non-convergence but has the same reproducability issue as in the univariate case
(+ slow):  

```{r bivariate_rstar, warning=FALSE}
rho <- 0.9
M <- 4
N <- 200
reps <- 20

r_functions <- c(r_star_gbm)
r_names <- c("R-star")
r_colors <- c(colors[2])

R_matrix <- repetitions_R(chains_func = (function() gen_bvnormal_chains(M, N, rho)), 
                         r_func = r_functions, 
                         r_names = r_names, 
                         reps = reps)

# plot_hist(R_matrix, colors = r_colors, nbreaks = 15, lim_y_axis = 10, plot_threshold = F)

plot_hist(R_matrix, colors = r_colors, bin_size = 0.04, 
          lim_y_axis = 10, vaxis_pos = 0.9,
          plot_threshold = F)

```

Finally the version of Vats and Knudson (2018) is not giving satisfying results
here:

```{r, warning=FALSE}
R_matrix_vats = repetitions_R(chains_func = (function() gen_bvnormal_chains(M, N, rho)), 
                             r_func = c(vats_multivariate_rhat), 
                             r_names = c("Vats Multivariate R-hat"), 
                             reps = reps)
print("Corresponding value of the R-hat version of Vats and Knudson (2018):")
print(R_matrix_vats[,1])

```

### Evolution of the different value of $\hat{R}$ with $\rho$

Except for $R^*$, we consider the mean of the 100 draws for the different 
versions of $\hat{R}$, and plot the evolution of this value with $\rho$:

```{r bivariate_rho, warning=FALSE}

rho_list = c(-0.99, -0.9, -0.7, -0.5, -0.2, 0, 0.2, 0.5, 0.7, 0.9, 0.99)

d <- 2
M <- 2
N <- 500
reps <- 10


r_functions <- c(rhat_infinity, brooks_multivariate_rhat)
r_names <- c("R-hat-infinity", "Brooks R-hat")

R_values <- c()
x_val <- c()
theoretical_R <- c()


for (rho in rho_list){
  R_val_rho = repetitions_R(chains_func = (function() gen_bvnormal_chains(M, N, rho)), 
                                    r_func = r_functions, 
                                    r_names = r_names, 
                                    reps = reps)
  R_values = rbind(R_values, R_val_rho)
  x_val = c(x_val, rep(rho, reps))
  
  sig_matrix <- (1-rho) * diag(2) + matrix(rho, nrow=2, ncol=2)
  dists <- c((function(q) mvtnorm::pmvnorm(upper = q, sigma = diag(2))[1]),
             (function(q) mvtnorm::pmvnorm(upper = q, sigma = sig_matrix)[1]))
  theoretical_R = c(theoretical_R, max_r_dist_bivariate(npoints = 200, 
                                                        xlim = c(-3,3), 
                                                        dists = dists))
}

# pdf(file = "/scratch/tmoins/Documents/Code_rhat/figure_article/rho_curve.pdf", width = 12, height = 7)
par(mar=c(4,5.5,2,2))
plot(x=x_val, y=R_values[,1],
     col = c(colors[3]), pch = 19,
     xlab = expression(rho["m"]), ylab = "",
     xaxs="i", yaxs="i", bty = "n",
     xaxt = "n", yaxt = "n",
     cex.lab = 2, cex.main = 2,
     lwd=9,
     xlim = c(-1.01, 1.02), ylim = c(0.999,1.155))
points(x=x_val, y=R_values[,2], 
       col = c(colors[4]), pch = 19,
       lwd=9)
lines(rho_list, theoretical_R, type="l", col = colors[2], lwd=6, lty = 2)
points(x=rho_list, y=theoretical_R, 
       col = c(colors[2]), pch = 19,
       lwd=9)
xlabels = c(1, 1.03, 1.06, 1.09, 1.12, 1.15)
ylabels = c(-1, -0.5, 0, 0.5, 1)
axis(1, labels=ylabels, at=ylabels, cex.axis=2, lwd=3, mgp = c(1,1,0))
axis(2, labels=xlabels, at=xlabels, cex.axis=2, mgp = c(1,0.5,0), lwd=2)
legend(x=0.6, y=1.14, col=c(colors[4], colors[3], colors[2]), pch=19, 
       legend = c(expression(italic(hat(R))),
                  expression(italic(hat(R)[infinity])),
                  expression(italic(R[infinity]))), cex=2)
# dev.off()
```

Remarks : 

- Our version is the most sensitive, and is more sensitive as $|\rho|$ increases
(which is great)
- The asymmetry can be link to the theoretical computations.
WARNING: in the case where $d=2$, the orders on the two bounds PQD and NQD are 
inverted (compared to the asymptotic behavior), and that is why here we are 
more sensitive when $\rho < 0$ (which is counter-intuitive)


```{r, include=FALSE, warning=FALSE}

d <- 2
M <- 2
N <- 500
reps <- 10

rho <- -0.9

r_functions <- c(coda_multivariate_rhat, brooks_multivariate_rhat)
r_names <- c("Brooks multivariate Rhat (Coda)", "Brooks multivariate Rhat (Lambert)")

R_matrix <- repetitions_R(chains_func = (function() gen_bvnormal_chains(M, N, rho)), 
                         r_func = r_functions, 
                         r_names = r_names, 
                         reps = reps)

# print(R_matrix)
# print(abs(R_matrix[,1]-R_matrix[,2]))
```


## General number of dimensions

### Generation of a covariance matrix

In the general case, we still compare $m-1$ chains that follows a multivariate
normal distribution with identity covariance matrix with one that has a given 
covariance matrix $\Sigma$. 

We generate $\Sigma$ from a sample of a Wishart distribtution $S$, and in order 
to keep the same margins for our chains, we transform $S$ to have ones on the
diagonal :
$$
\Sigma = D^{-1/2} S D^{-1/2}, 
\quad \text{ with } \quad 
D = \text{Diag}\left(s_{1,1}, \ldots, s_{d,d}\right)
$$

From the theoretical part, we know that the value given by $\hat{R}_\infty$ 
depends on the direction of dependence.
Thus, we can either fix $\Sigma$ and study the behaviour of the different 
$\hat{R}$, or sample a new $\Sigma$ for each draw of $\hat{R}$.

```{r, warning=FALSE}

generate_sigma_matrix <- function(d){
  S <- rWishart(1, d, diag(d))[,,1]
  S <- diag(1/sqrt(diag(S))) %*% S %*% diag(1/sqrt(diag(S)))
  return (S)
}

gen_mvnormal_chains <- function(M, N, d, S=NULL){
  if (is.null(S)){
    S <- generate_sigma_matrix(d)
  }
  return (array(c(mvrnorm((M-1)*N, mu = rep(0, d), Sigma = diag(d)), 
                  mvrnorm(N, mu = rep(0, d), Sigma = S)), 
                c(N,d,M)))
}

```

First, we start by considering $d=5$ and study the behavior of $\hat{R}_\infty$
and the multivariate $\hat{R}$ of Brooks and Gelman (1998) on a given $\Sigma$
for the last chain:

```{r multivariate_hist1, warning=FALSE}

d <- 5

M <- 4
N <- 100
reps <- 100

r_functions <- c(brooks_multivariate_rhat, rhat_infinity)
r_names <- c("Brooks Multivariate R-hat", "R-hat-infinity")
r_colors <- c(colors[3], colors[4])

S <- generate_sigma_matrix(d)
print(S)

R_matrix <- repetitions_R(chains_func = (function() gen_mvnormal_chains(M, N, d, S)), 
                         r_func = r_functions, 
                         r_names = r_names, 
                         reps = reps)

# plot_hist(R_matrix, colors = r_colors, nbreaks = 20, lim_y_axis = 30, plot_threshold = F)
plot_hist(R_matrix, colors = r_colors, bin_size = 0.002, 
          lim_y_axis = 25, vaxis_pos = 1,
          plot_threshold = F)

```

Doing the same experiment but with a new $\Sigma$ on each repetitions will
result on a histogram more spread for $\hat{R}_\infty$:

```{r multivariate_hist2, warning=FALSE}


d <- 5

M <- 4
N <- 100
reps <- 100

r_functions <- c(brooks_multivariate_rhat, rhat_infinity)
r_names <- c("Brooks Multivariate R-hat", "R-hat-infinity")
r_colors <- c(colors[3], colors[4])

R_matrix <- repetitions_R(chains_func = (function() gen_mvnormal_chains(M, N, d)), 
                         r_func = r_functions, 
                         r_names = r_names, 
                         reps = reps)

# plot_hist(R_matrix, colors = r_colors, nbreaks = 20, lim_y_axis = 30, plot_threshold = F)
xlabels = c(1, 1.02, 1.04, 1.06, 1.08)
plot_hist(R_matrix, colors = r_colors, 
          xlabels = xlabels, bin_size = 0.002, 
          lim_y_axis = 20, vaxis_pos = 0.997,
          plot_threshold = F)

```

\textit{Le problème en grande dimension : aucun point n'a toutes ces coordonnées plus petite que les autres ! Déjà lorsque d = 2, le problème est déjà présent avec plusieurs points produisant des NaN values. Exemple avec d=2, N=100, M=2:

-0.9640542 -2.8385044
-0.7183567 -3.3431770
-2.579755 -1.719382
-1.762459 -1.977133

Tout ces points sont des """minimas""", à l'inverse du cas d=1 où 1 seul est possible.}

To avoid this issue, $\hat{R}_\infty$ can be computed on all directions, 
which means if $d=3$ that $\hat{R}$ is computed on 
$$
  I\left(\theta_1^{(n,m)} \leq x_1, \theta_2^{(n,m)} \leq x_2, \theta_3^{(n,m)} \leq x_3\right)\\
  I\left(\theta_1^{(n,m)} \geq x_1, \theta_2^{(n,m)} \leq x_2, \theta_3^{(n,m)} \leq x_3\right)\\
  I\left(\theta_1^{(n,m)} \leq x_1, \theta_2^{(n,m)} \geq x_2, \theta_3^{(n,m)} \leq x_3\right)\\
  I\left(\theta_1^{(n,m)} \geq x_1, \theta_2^{(n,m)} \geq x_2, \theta_3^{(n,m)} \leq x_3\right)\\
$$
The other half of the possibilities is avoided by symmetry. 
Then, we can define $\hat{R}^{(\text{max})}_\infty$ as the maximum of all the 
$\hat{R}_\infty$ computed on all directions.
Despite a unreasonable computational cost (with $2^{d-1}$ $\hat{R}_\infty$ 
to compute), this solutions seems satisfying in small dimension:

```{r r_dimensions, warning=FALSE}

# d_max = 6
# 
# M = 4
# N = 100
# reps = 200
# 
# 
# r_functions = c(rhat_infinity, rhat_infinity_max_directions)
# r_names = c("One Direction", "All Directions")
# 
# R_values = repetitions_R(chains_func = (function() gen_mvnormal_chains(M, N, 2)), 
#                          r_func = r_functions, 
#                          r_names = r_names, 
#                          reps = reps)
# 
# 
# R_array = array(c(R_values[,1], rep("d = 2", reps), rep("1 Dir", reps)), dim = c(reps,3))
# R_array = rbind(R_array, array(c(R_values[,2], rep("d = 2", reps), rep("All Dir", reps)), dim = c(reps,3)))
# colnames(R_array) = c("Values", "Dim", "Nb_Dir")
# 
# for (d in 3:d_max){
#   R_val_d = repetitions_R(chains_func = (function() gen_mvnormal_chains(M, N, d)), 
#                              r_func = r_functions, 
#                              r_names = r_names, 
#                              reps = reps)
#   d_names = c(paste("d =", d))
#   R_array = rbind(R_array, array(c(R_val_d[,1], rep(d_names, reps), rep("1 Dir", reps)), dim = c(reps,3)))
#   R_array = rbind(R_array, array(c(R_val_d[,2], rep(d_names, reps), rep("All Dir", reps)), dim = c(reps,3)))
# 
# }
# 
# R_df = as.data.frame(R_array)
# R_df$Values = as.numeric(R_df$Values)
# 
# # pdf(file = "/scratch/tmoins/Documents/Code_rhat/figure_article/hist_directions.pdf", width = 18, height = 8)
# ggplot(R_df, aes(x = Values, y = as.factor(Dim), fill = Nb_Dir)) + 
#   geom_density_ridges(stat = "binline", bins= 40,
#                       scale = 1, alpha = 0.7) +
#   scale_y_discrete(limits = R_df$Dim) +
#   ylab("Dimensions") + labs(fill = element_blank()) +
#   scale_x_continuous(limits = c(1,1.14),
#                      n.breaks = 8) +
#   scale_fill_manual(values=c("#32a632", "#4287f5")) +
#   theme_ridges(font_size=22)
# # dev.off()

```

(WARNING: $d>6$ the chunk is very dangerous to run!!)

## Logistic regression

As a practical bayesian example, we consider hierarchical logistic regression using 
the following model:

```{r, warning=FALSE}
writeLines(readLines(paste(stan_folder, "logit_reg.stan", sep="")))
```

We run Stan using the dataset \texttt{logit} available in the package \texttt{mcmc}.

```{r, warning=FALSE}
data("logit")
d <- 4
logit_data<- read_rdump(paste(stan_folder, "logit_reg.data.R", sep=""))
fit_logit <- stan(file = paste(stan_folder, "logit_reg.stan", sep=""),
                  data=logit_data, seed = 12,
                  iter=400, chains=4, refresh=0)
mon <- monitor(fit_logit)

chains <- as.array(fit_logit)
dim_chains <- dim(chains)

rhat_inf <- c()

for (i in 1:dim_chains[3]) {
  chains_i <- chains[, , i]
  rhat_inf_i <- rhat_infinity(chains_i, max_nb_points = "ALL")
  rhat_inf <- c(rhat_inf, rhat_inf_i)
}

mon[["Rhat_inf"]] <- rhat_inf

mon
summary(mon[1:d, 'Rhat'])
summary(mon[1:d, 'Rhat_inf'])

mon[d+1, "Rhat_inf"]

perm_chains <- aperm(chains, c(1,3,2))
rhat_infinity_max_directions(perm_chains)
```

```{r, warning=FALSE}
reps <- 50

gen_beta_logit <- function(i){
  # fit_logit <- stan(file = paste(stan_folder, "logit_reg.stan", sep=""),
                    # data=logit_data,
                    # iter=400, chains=4, refresh=0)
  # saveRDS(fit_logit, paste(stan_folder, "logit/logit_reg_", i, ".rds", sep=""))
  fit_logit <- readRDS(paste(stan_folder, "logit/logit_reg_", i, ".rds", sep=""))
  mon <- monitor(fit_logit)
  
  chains <- as.array(fit_logit)
  
  return (aperm(chains, c(1,3,2)))
}

r_functions = c(brooks_multivariate_rhat, rhat_infinity_max_directions)
r_names = c("Brooks Multivariate R-hat", "max-R-hat")
r_colors = c(colors[4], colors[5])

R_matrix_logit <- repetitions_R(chains_func = gen_beta_logit, 
                                 r_func = r_functions, 
                                 r_names = r_names, 
                                 reps = reps, saved_model = TRUE)

# pdf(file = "/scratch/tmoins/Documents/Code_rhat/figure_article/hist_logit.pdf", width = 12, height = 7)
plot_hist(R_matrix_logit, colors = r_colors, 
          vaxis_pos = 0.998,
          bin_size = 0.003, lim_y_axis = 25)
# dev.off()
```


```{r, warning=FALSE}

gen_beta_logit_lp <- function(i){
  fit_logit <- readRDS(paste(stan_folder, "logit/logit_reg_", i, ".rds", sep=""))
  mon <- monitor(fit_logit)
  chains <- as.array(fit_logit)
  return (chains[, , 5])
}

r_functions <- c(rhat_infinity)
r_names <- c("R-hat-infinity")
r_colors <- c(colors[3])

R_matrix_logit2 <- repetitions_R(chains_func = gen_beta_logit_lp, 
                                 r_func = r_functions, 
                                 r_names = r_names, 
                                 reps = reps, saved_model = TRUE)

xlabels = c(1, 1.005, 1.01, 1.015, 1.02, 1.025)
# pdf(file = "/scratch/tmoins/Documents/Code_rhat/figure_article/hist_logit2.pdf", width = 12, height = 7)
plot_hist(R_matrix_logit2, colors = r_colors,
          xlabels = xlabels,
          bin_size = 0.002, lim_y_axis = 25)
# dev.off()
```


# Wide multivariate normal

```{r, warning=FALSE}
writeLines(readLines(paste(stan_folder, "mvn.stan", sep="")))
```

```{r, warning=FALSE}
generate_sigma_matrix <- function(d){
  S = rWishart(1, d, diag(d))[,,1]
  S = diag(1/sqrt(diag(S))) %*% S %*% diag(1/sqrt(diag(S)))
  return (S)
}

d <- 8
A <- generate_sigma_matrix(d)
fit <- stan(file = paste(stan_folder, "mvn.stan", sep=""), data=list(N=d, A=A), iter=400, chains=4, refresh = 0)
mon <- monitor(fit)

chains <- as.array(fit)
dim_chains <- dim(chains)

rhat_inf <- c()

for (i in 1:dim_chains[3]) {
  chains_i <- chains[, , i]
  rhat_inf_i <- rhat_infinity(chains_i, max_nb_points = "ALL")
  rhat_inf <- c(rhat_inf, rhat_inf_i)
}

mon[["Rhat_inf"]] <- rhat_inf
# summary(mon[1:d, 'Rhat_inf'])
mon
# mon[d+1, "Rhat_inf"]
```


```{r, warning=FALSE}
which_max_rhat_inf <- which.max(mon[1:d, 'Rhat_inf'])

simulated_rhat = all_local_rhat(chains[ , , which_max_rhat_inf], max_nb_points = "ALL")
plot_local_r(chains[ , , which_max_rhat_inf], simulated_rhat,
             col=c(colors[3], colors[2]),
             xlim = c(-3,3), ylim=c(0.997,1.02), title ="")
```


```{r, warning=FALSE}
simulated_rhat = all_local_rhat(chains[ , , d+1], max_nb_points = "ALL")
plot_local_r(chains[ , , d+1], simulated_rhat,
             col=c(colors[3], colors[2]),
             xlim = c(min(chains), 0), ylim=c(0.997,1.02), title ="")
```


```{r, warning=FALSE}
perm_chains <- aperm(chains, c(1,3,2))
rhat_infinity_max_directions(perm_chains)
```


```{r, warning=FALSE}
d <- 8
A <- generate_sigma_matrix(d)
fit_mvn <- stan(file = paste(stan_folder, "mvn.stan", sep=""), data=list(N=d, A=A), iter=800, chains=4, refresh = 0)
mon <- monitor(fit_mvn)

chains <- as.array(fit_mvn)
dim_chains <- dim(chains)

rhat_inf <- c()

for (i in 1:dim_chains[3]) {
  chains_i <- chains[, , i]
  rhat_inf_i <- rhat_infinity(chains_i, max_nb_points = "ALL")
  rhat_inf <- c(rhat_inf, rhat_inf_i)
}

mon[["Rhat_inf"]] <- rhat_inf
mon
summary(mon[1:d, 'Rhat_inf'])

# mon[d+1, "Rhat_inf"]
```


```{r, warning=FALSE}
perm_chains <- aperm(chains, c(1,3,2))
rhat_infinity_max_directions(perm_chains)
```



```{r, warning=FALSE}
d <- 8
A <- generate_sigma_matrix(d)
fit <- stan(file = paste(stan_folder, "mvn.stan", sep=""), data=list(N=d, A=A), iter=100, chains=4, refresh = 0)
mon <- monitor(fit)

chains <- as.array(fit)
dim_chains <- dim(chains)

rhat_inf <- c()

for (i in 1:dim_chains[3]) {
  chains_i <- chains[, , i]
  rhat_inf_i <- rhat_infinity(chains_i, max_nb_points = "ALL")
  rhat_inf <- c(rhat_inf, rhat_inf_i)
}

mon[["Rhat_inf"]] <- rhat_inf
# summary(mon[1:d, 'Rhat_inf'])
mon
# mon[d+1, "Rhat_inf"]
perm_chains <- aperm(chains, c(1,3,2))
vats_multivariate_rhat(perm_chains)
```

